---
title: "Project1"
author: "LeTicia Cancel"
date: "9/18/2020"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 2
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# Libraries
#library(tidyverse)
library(RCurl)
library(dplyr)
library(stringr)
library(tidyr)
```

## Import Tournament Data

The tournament text file is saved in GitHub. The data is saved to the table t_df and the columns are renamed.
```{r}
#get URL and import data from GitHub
site <- getURL("https://raw.githubusercontent.com/ltcancel/Project1_Data607_F20/master/tounamentinfo.txt")
df <- read.delim(text = site, header = FALSE,sep = "|")

#create a tibble with the imported data
t_df <- as_tibble(df)

#rename columns
cnames <- c('Pair','PlayerName','TotalPts','Round1','Round2','Round3','Round4','Round5','Round6','Round7','drop')
colnames(t_df) <- cnames

head(t_df)
```

## Data Cleanup

After the data is saved to a table we need to remove all unnecessary information. 
```{r}
#get all rows with the pair number and player name
df_filter1 <- t_df %>%
  filter(str_detect(Pair, regex("\\d"))) %>%
  filter(!str_detect(str_to_lower(TotalPts), "n"))

#get all rows that include the other player information
df_filter2 <- t_df %>%
  filter(str_detect(str_to_lower(Pair), "o|m")) %>%
  filter(!str_detect(str_to_lower(TotalPts), "p"))

#combine both tables
df_combined <- cbind(df_filter1[,1:10], df_filter2[,1:2])

cnames <- c("Playernum","PlayerName","TotalPts",'Round1','Round2','Round3','Round4','Round5','Round6','Round7','State','ratings')

#rename columns again
colnames(df_combined) <- cnames

head(df_combined)

```

Some columns contain multiple types of information, so those will need to be split into their own columns.
```{r warning=FALSE}
#Separate column that holds pre and post ratings
df_combined <- df_combined %>%
  separate(ratings, c("drop", "keep"), ": ")

df_combined <- df_combined %>%
  separate(keep, c("Pre-Rating", "Post-Rating"), "->")

df_combined <- df_combined %>%
  separate(`Pre-Rating`, c("Pre-Rating", "drop2"), "P")

#select only the columns needed after splitting the data
df_combined <- df_combined %>%
  select(1:11,13,15)

head(df_combined)
```
Additional splitting is needed for the columns that hold the player number of each of their opponents.
```{r}
#split round 1
df_combined <- df_combined %>%
  separate(Round1, c("R1WL","R1Player"))

#split round 2
df_combined <- df_combined %>%
  separate(Round2,c("R2WL","R2Player"))

#split round 3
df_combined <- df_combined %>%
  separate(Round3,c("R3WL","R3Player"))

#split round 4
df_combined <- df_combined %>%
  separate(Round4,c("R4WL","R4Player"))

#split round 5
df_combined <- df_combined %>%
  separate(Round5,c("R5WL","R5Player"))

#split round 6
df_combined <- df_combined %>%
  separate(Round6,c("R6WL","R6Player"))

#split round 7
df_combined <- df_combined %>%
  separate(Round7,c("R7WL","R7Player"))

#Select only the columns needed after spliting the data
df_combined <- df_combined %>%
  select(1:3,5,7,9,11,13,15,17:19)

head(df_combined)
```

## Computing Averages

Here we need to do some additional work to get the average of all opponent's pre-rating.
```{r}
#Create a new table that only holds the player number and their pre-ratings
ratings <- df_combined %>%
  select(Playernum, `Pre-Rating`)

head(ratings)
```

When I tried to compute the average for each row I realized that it was not working because the columns needed for the equation were all stored as "chr" data types. Here I convert all columns needed for joining and computing to an integer.
```{r}
#convert data to an integer so it can be matched
ratings$Playernum <- as.integer(ratings$Playernum)

#loop through the main table to convert all player number columns into an int
for(i in 4:10){
  df_combined[,i] <- as.integer(df_combined[,i])
}

#join the ratings table and the df_combined table so we can see the pre-rating of each opponent
df_combined <- df_combined %>%
  inner_join(ratings, by = c("R1Player" = "Playernum")) %>%
  inner_join(ratings, by = c("R2Player" = "Playernum")) %>%
  inner_join(ratings, by = c("R3Player" = "Playernum")) %>%
  inner_join(ratings, by = c("R4Player" = "Playernum")) %>%
  inner_join(ratings, by = c("R5Player" = "Playernum")) %>%
  inner_join(ratings, by = c("R6Player" = "Playernum")) %>%
  inner_join(ratings, by = c("R7Player" = "Playernum"))

#convert new columns added from the join above to an integer so they can be averaged
for(i in 12:19){
  df_combined[,i] <- as.integer(df_combined[,i])
}


#compute the average of the pre-ratings for each row
df_combined <- mutate(df_combined, (rowSums(df_combined[,12:19], na.rm = TRUE))/7)

head(df_combined)
```

## Export to CSV

Now we create the final table and export it to a csv file.
```{r}
#create the final tables only using the columns that need to be exported
final_table <- df_combined %>%
  select(2,3,11,12,20)

#remane final table
names <- c("PlayerName", "TotalPts", "State", "Pre-Rating", "OppAvgRating")
colnames(final_table) <- names

head(final_table)
```

```{r}
#write to a csv file
write.csv(final_table,"FinalData.csv")
```

